---
title: "Baltimore Liquor"
author: "BODF"
date: "12/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Current date and time (format = Weekday Month Day Hour:Min:Sec Year):
```{r date, echo = TRUE}
date()
```
# Liquor licenses in Baltimore City  
The Baltimore open data website allows us to access a wide range of data sets 
related to the operation of a city. I decided to download their liqour license 
data and plot it against the locations of the licenses (where can I find the 
most beer?).  

The data are broken down by license end date (I chose 2018 for relevancy), 
name of the establishment, liquor license type, and location. The names are 
plotted as popup messages as you select individual markers. The liquor license 
type dictates the color, note the types are as follows:  

* B = Beer
* BW = Beer and Wine
* BWL = Beer, Wine, and Liquor
* AE = Adult Entertainment (unclear why Baltimore makes this distinction)

The data markers are clustered in the map, so the liquor license type (color) 
will not be obvious until you zoom in on individual licenses.

```{r, echo = FALSE}
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(stringr))

# Import the data, filter out only data with a license that goes thru 2018,
# then select out just the columns we need: TradeName, SubClass, `Location 1`
# make a map of these vars with color dictated by Subclass and popup by TradeName
setwd("~/Documents/Classes/DataSciCourse/")
suppressMessages(data <- read_csv("Liquor_Licenses.csv", na = c("", "NA", "N/A")) %>%
        mutate(LicenseEndDate = mdy(LicenseEndDate)) %>%
        filter(LicenseEndDate >= mdy("01/01/2018"))%>% # grab only the pubs still open
        select(SubClass, TradeName, `Location 1`) %>%
        rename(Location = `Location 1`))

# below code extracts any parenthetical expressions in "Location". Some of this 
# is lat by longitude, some is not: so I include a logical test. If passed, the 
# string is trimmed and latitude + longitude are extracted, separately
index <- 1
coord_nums <- data.frame(lng = NA, lat = NA)
suppressWarnings(for(element in data$Location){
    buzz <- strsplit(str_extract(element, "\\(.*\\)"), "")[[1]]
    if(!is.na(as.numeric(buzz[2]))){
        buffer <- buzz[-c(1,length(buzz))]
        id <- which(buffer==",")
        latit <- as.numeric(paste(buffer[1:(id-1)], collapse=''))
        longit <- as.numeric(paste(buffer[(id+1):length(buffer)], collapse=''))
        coord_nums[index, 1] <- longit
        coord_nums[index, 2] <- latit
        index <- index+1
    }
})

# grab a subsettable logical list and trim data using it
value <- logical()
suppressWarnings(for(element in data$Location){
    buzz <- strsplit(str_extract(element, "\\(.*\\)"), "")[[1]]
    if(!is.na(as.numeric(buzz[2]))){
        value <- c(value, TRUE)
    } else {value <- c(value, FALSE)}
})
data <- data[value,]

# now replace Location with lat x long
data <- data %>% 
        select(-Location) %>% 
        mutate(lat = coord_nums$lat, lng = coord_nums$lng)
rm(coord_nums) # free up memory

# assign colors based on subClass
coloring <- character()
for(element in data$SubClass){
    if(element == "AE"){
        coloring <- c(coloring, "#ca0020")
    } else if(element == "B"){
        coloring <- c(coloring, "#74c476")
    } else if(element == "BW"){
        coloring <- c(coloring, "#969696")
    } else if(element == "BWL"){
        coloring <- c(coloring, "#2171b5")
    }
}

# Make the Map!
data <- as.data.frame(data)
suppressMessages(data %>% 
        leaflet() %>% 
        addTiles() %>% 
        addCircleMarkers(color = coloring, 
                   clusterOptions = markerClusterOptions(), 
                   popup = data$TradeName) %>%
        addLegend(labels = c("AE", "B", "BW", "BWL"), 
                  colors=c("#ca0020", "#74c476", "#969696", "#2171b5")))
```